---
title: "Project"
output: pdf_document
date: "2023-11-21"
---

Goal: Optimizing Policy Revenue Predictions Using Calibrated Probabilistic Classifiers Under Class Imbalance

```{r}
library(caret)
library(glmnet)
library(randomForest)
library(verification)
library(rms)
library(zoo)
library(gbm)
library(pROC)
library(dplyr)
library(pROC)

```

```{r}
df = read.csv('aug_train.csv')
#Dropping the id column since we already have dataframe indices as their unique ID
df = df[, !names(df) %in% 'id']
head(df)
```

Description of features:

Gender: Describes the gender of a customer
Age: Describes the age of a customer
Driving_license: An binary feature which takes the value of 1 when the customer possess a driving license
Region_code: The region a customer belongs to (Ranging from 0 to 52)
Previously_Insured: A binary feature which takes the value of 1 when the customer possessed a vehicle insurance before.
Vehicle_Age: Describes the age of a customer's vehicle
Vehicle_Damage: A feature taking the values of either Yes or No according to if a customer's vehicle is damaged or not
Annual_Premium: Amount of premium a customer will pay if they buy a vehicle insurance (Currency undefined)
Policy_Sales_Channel: The unique identifier of referral agency
Vintage: The number of days the customer has been associated with the company
Response: A binary feature which takes the value of 1 when the customer purchased a vehicle insurance.
```{r}
#Making factors of some categorical variables
df$Region_Code <- as.factor(df$Region_Code)
df$Policy_Sales_Channel <- as.factor(df$Policy_Sales_Channel)
df$Previously_Insured <- as.factor(df$Previously_Insured)
df$Gender <- as.factor(df$Gender)
df$Driving_License <- as.factor(df$Driving_License)
df$Vehicle_Age <- as.factor(df$Vehicle_Age)
df$Vehicle_Damage <- as.factor(df$Vehicle_Damage)
```

### Test Train Split

```{r}
set.seed(1010137629)
index <- createDataPartition(df$Response, p = 0.8, list = FALSE)
train <- df[index, ]
test <- df[-index, ]
```

### Data Preprocessing


```{r}
#dummy_transform <- dummyVars(" ~ .", data = preprocessed_df)
dummy_transform <- dummyVars(" ~ .", data = train)

# Apply the dummy transformation to the data
#preprocessed_df <- data.frame(predict(dummy_transform, newdata = preprocessed_df))
preprocessed_df <- data.frame(predict(dummy_transform, newdata = train))
```

```{r}
#Dropping features to avoid Variable Trap
# preprocessed_df = preprocessed_df[, !names(preprocessed_df) %in% c('GenderMale', 'Vehicle_Age_morethan2', 'Vehicle_DamageYes', 'Region_Code.0', 'Policy_Sales_Channel.1')]
```

```{r}
y <- preprocessed_df$Response
#y <- as.factor(y)
X <- as.matrix(preprocessed_df[, !names(preprocessed_df) %in% 'Response'])
```

### Modelling

```{r}
set.seed(1010137629)
rf_model <- randomForest(X, as.factor(y), ntree = 30, do.trace=TRUE)
```
```{r}
predictions <- predict(rf_model, newx = X, type = "prob")
#predictions <- na.fill(as.matrix(predictions), fill=0)
predictions <- predictions[,2]
```


```{r}
options(scipen = 999)
result_df <- data.frame(Actual = y, Predictions = predictions)
#result_df$Actual <- as.numeric(result_df$Actual)
#result_df$Predictions <- as.numeric(result_df$Predictions)
result_df$Brier_Score <- (result_df$Actual - result_df$Predictions)^2
#0 being the majority class
result_df$Baseline_Brier_Score <- (result_df$Actual - 0)^2
result_df
```

```{r}
cat('Brier Score: ', mean(result_df$Brier_Score), '\n', 'Baseline Brier Score: ',
    mean(result_df$Baseline_Brier_Score))
```

```{r}
# Create a reliability diagram
reliability.plot(verify(result_df$Actual, result_df$Predictions))
```

```{r}
library(pROC)

a <- result_df
a$Actual <- as.numeric(a$Actual)
roc(a$Actual, a$Prediction)
```
```{r}
roc_curve_obj <- roc(a$Actual, a$Prediction, direction = "auto")
auc_pr <- auc(roc_curve_obj, curve = TRUE)
plot(roc_curve_obj, col = "blue", main = "Precision-Recall Curve")
```

### Testing

```{r}
dummy_transform <- dummyVars(" ~ .", data = test)

# Apply the dummy transformation to the data
#preprocessed_df <- data.frame(predict(dummy_transform, newdata = preprocessed_df))
preprocessed_df_test <- data.frame(predict(dummy_transform, newdata = test))
```


```{r}
y_test <- preprocessed_df_test$Response
#y <- as.factor(y)
X_test <- as.matrix(preprocessed_df_test[, !names(preprocessed_df_test) %in% 'Response'])
```

```{r}
predictions_test <- predict(rf_model,X_test, type = "prob")[,2]

```

```{r}
result_df_test <- data.frame(Actual = y_test, Predictions = predictions_test)
#result_df$Actual <- as.numeric(result_df$Actual)
#result_df$Predictions <- as.numeric(result_df$Predictions)
result_df_test$Brier_Score <- (result_df_test$Actual - result_df_test$Predictions)^2
#0 being the majority class
result_df_test$Baseline_Brier_Score <- (result_df_test$Actual - 0)^2
result_df_test
```
```{r}
reliability.plot(verify(result_df_test$Actual, result_df_test$Predictions, nbins=20))
```
```{r}
cat('Brier Score: ', mean(result_df_test$Brier_Score), '\n', 'Baseline Brier Score: ',
    mean(result_df_test$Baseline_Brier_Score))
```